Supporting Information for "Imperfect slope measurements drive overestimation in geometric models of lake depth"

### 2021-08-12 manuscript/si_template_2019.tex
linting

% source at tables/01_predictors.Rmd
\begin{table}
\caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.} \label{tableS1}
\centering
\setlength\tabcolsep{1.5pt} % default value: 6pt
\begin{tabular}{lllll}
\hline
Variable & Median & Q25 & Q75 & n\\
\hline
Max depth (m) & 8.2 (7) & 4.6 (3.7) & 14 (12) & 4820 (17700)\\
\textbf{Area (ha)} & 55 (33) & 21 (11) & 140 (100) & 4820 (17700)\\
\textbf{Island area (ha)} & 0 (0) & 0 (0) & 0.18 (0.076) & 4820 (17700)\\
\textbf{Perimeter (m)} & 4400 (3500) & 2500 (1800) & 8100 (7300) & 4820 (17700)\\
\textbf{Shoreline development} & 1.7 (1.7) & 1.4 (1.4) & 2.1 (2.2) & 4820 (17700)\\
\textbf{Elevation (m)} & 300 (340) & 180 (210) & 400 (460) & 4820 (17700)\\
\textbf{Watershed-lake ratio} & 7.8 (10) & 3.8 (4.4) & 17 (29) & 4820 (17700)\\
\textbf{Deepest point distance (m)} & 180 (-) & 110 (-) & 290 (-) & 4820 (-)\\
Mean deepest point distance (m) & 140 (-) & 87 (-) & 230 (-) & 4820 (-)\\
\textbf{Visual center distance (m)} & 240 (-) & 160 (-) & 390 (-) & 4820 (-)\\
Inlake slope (m/m) & 0.05 (-) & 0.02 (-) & 0.08 (-) & 4820 (-)\\
Inlake slope online (m/m) & 0.06 (-) & 0.03 (-) & 0.14 (-) & 4800 (-)\\
Inlake slopes (m/m) & 0.06 (-) & 0.03 (-) & 0.1 (-) & 4820 (-)\\
Inlake slopes online (m/m) & 0.07 (-) & 0.03 (-) & 0.15 (-) & 4800 (-)\\
Mean inlake slope (m/m) & 0.04 (-) & 0.02 (-) & 0.09 (-) & 4820 (-)\\
Nearshore mean slope (m/m) & 0.08 (-) & 0.05 (-) & 0.11 (-) & 4820 (-)\\
Nearshore slope online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4590 (-)\\
Nearshore slopes online (m/m) & 0.08 (-) & 0.04 (-) & 0.13 (-) & 4540 (-)\\
\hline
\end{tabular}
\end{table}

\begin{table}
\caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination, MAPE = mean absolute percent error) for the proxy - proxy combination of geometry metrics (see main text Table 1). Each row shows model metrics when proxy and "true" measures are calculated with slight differences from the default (bolded) used in the main text. \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points. \texttt{slope\_pnts} is the average slope (i.e. \texttt{slope\_pnt}) of all points at maximum depth. \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope). \texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth. \texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}  \label{tableS2}
\centering
\setlength\tabcolsep{1.5pt} % default value: 6pt
\begin{tabular}{lllllc}
\hline
Inlake slope & Nearshore slope & Inlake distance & RMSE & $R^2$ & MAPE\\
\hline
slope\_pnts & slope\_mean & dists\_deepest & 6.2 m & 0.38 & 58 \%\\
\textbf{slope\_pnt} & \textbf{slope\_mean} & \textbf{dist\_deepest} & \textbf{6.4 m} & \textbf{0.35} & \textbf{59 \%}\\
slope\_pnts & slopes\_online\_mean & dist\_deepest & 6.4 m & 0.32 & 61 \%\\
slope\_online\_mean & slope\_mean & dists\_deepest & 6.5 m & 0.41 & 63 \%\\
slope\_pnts & slope\_mean & dist\_deepest & 6.7 m & 0.44 & 58 \%\\
slope\_online\_mean & slope\_mean & dist\_deepest & 6.7 m & 0.36 & 59 \%\\
slope\_online\_mean & slopes\_online\_mean & dist\_deepest & 6.7 m & 0.32 & 66 \%\\
slope\_mean & slope\_mean & dists\_deepest & 6.8 m & 0.36 & 59 \%\\
slope\_pnt & slopes\_online\_mean & dists\_deepest & 6.8 m & 0.25 & 73 \%\\
slope\_pnts & slope\_online\_mean & dist\_deepest & 6.9 m & 0.3 & 71 \%\\
slope\_online\_mean & slope\_online\_mean & dist\_deepest & 6.9 m & 0.32 & 68 \%\\
slope\_online\_mean & slope\_online\_mean & dists\_deepest & 6.9 m & 0.33 & 65 \%\\
slope\_mean & slopes\_online\_mean & dists\_deepest & 7 m & 0.24 & 65 \%\\
slope\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.4 & 64 \%\\
slopes\_online\_mean & slope\_mean & dist\_deepest & 7.1 m & 0.37 & 56 \%\\
slope\_mean & slope\_online\_mean & dist\_deepest & 7.1 m & 0.3 & 69 \%\\
slopes\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.2 m & 0.32 & 63 \%\\
slopes\_online\_mean & slopes\_online\_mean & dist\_deepest & 7.3 m & 0.25 & 64 \%\\
slope\_pnt & slope\_mean & dists\_deepest & 7.3 m & 0.35 & 61 \%\\
slopes\_online\_mean & slope\_mean & dists\_deepest & 7.3 m & 0.36 & 60 \%\\
slope\_online\_mean & slopes\_online\_mean & dists\_deepest & 7.3 m & 0.29 & 58 \%\\
slope\_pnts & slope\_online\_mean & dists\_deepest & 7.4 m & 0.27 & 64 \%\\
slopes\_online\_mean & slope\_online\_mean & dists\_deepest & 7.4 m & 0.33 & 67 \%\\
slopes\_online\_mean & slope\_online\_mean & dist\_deepest & 7.5 m & 0.26 & 61 \%\\
slope\_pnt & slopes\_online\_mean & dist\_deepest & 7.5 m & 0.33 & 69 \%\\
slope\_mean & slopes\_online\_mean & dist\_deepest & 7.6 m & 0.26 & 64 \%\\
slope\_pnt & slope\_online\_mean & dist\_deepest & 7.7 m & 0.27 & 68 \%\\
slope\_pnts & slopes\_online\_mean & dists\_deepest & 7.8 m & 0.3 & 65 \%\\
slope\_pnt & slope\_online\_mean & dists\_deepest & 7.9 m & 0.27 & 67 \%\\
slope\_mean & slope\_online\_mean & dists\_deepest & 7.9 m & 0.31 & 60 \%\\
\hline
\end{tabular}
\end{table}

>>>>

% source at tables/01_predictors.Rmd
\begin{table}
      \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.} \label{tableS1}
      \centering
      \setlength\tabcolsep{1.5pt} % default value: 6pt
      \begin{tabular}{lllll}
            \hline
            Variable                            & Median      & Q25         & Q75          & n            \\
            \hline
            Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
            \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
            \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
            \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
            \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
            \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
            \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
            \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
            Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
            \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
            Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
            Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
            Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
            Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
            Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
            Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
            Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
            Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
            \hline
      \end{tabular}
\end{table}

\begin{table}
      \caption{Model fit and predictive accuracy metrics (RMSE = root mean square error, $R^2$ = coefficient of determination, MAPE = mean absolute percent error) for the proxy - proxy combination of geometry metrics (see main text Table 1). Each row shows model metrics when proxy and "true" measures are calculated with slight differences from the default (bolded) used in the main text. \texttt{slope\_mean} is the mean slope of all inlake or nearshore buffer points. \texttt{slope\_pnts} is the average slope (i.e. \texttt{slope\_pnt}) of all points at maximum depth. \texttt{slope\_online\_mean} is the mean pixel-to-pixel slope of each pixel lying on a straight line either from the single deepest point to the lake shoreline (in the case of inlake slope) or from the lake shoreline point extending to the buffer exterior (in the case of nearshore slope). \texttt{slopes\_online\_mean} is the same as \texttt{slope\_online\_mean} except it uses all inlake points at maximum depth. \texttt{dists\_deepest} is the same as \texttt{dist\_deepest} except distance is calculated for all points at maximum depth.}  \label{tableS2}
      \centering
      \setlength\tabcolsep{1.5pt} % default value: 6pt
      \begin{tabular}{lllllc}
            \hline
            Inlake slope         & Nearshore slope      & Inlake distance        & RMSE           & $R^2$         & MAPE           \\
            \hline
            slope\_pnts          & slope\_mean          & dists\_deepest         & 6.2 m          & 0.38          & 58 \%          \\
            \textbf{slope\_pnt}  & \textbf{slope\_mean} & \textbf{dist\_deepest} & \textbf{6.4 m} & \textbf{0.35} & \textbf{59 \%} \\
            slope\_pnts          & slopes\_online\_mean & dist\_deepest          & 6.4 m          & 0.32          & 61 \%          \\
            slope\_online\_mean  & slope\_mean          & dists\_deepest         & 6.5 m          & 0.41          & 63 \%          \\
            slope\_pnts          & slope\_mean          & dist\_deepest          & 6.7 m          & 0.44          & 58 \%          \\
            slope\_online\_mean  & slope\_mean          & dist\_deepest          & 6.7 m          & 0.36          & 59 \%          \\
            slope\_online\_mean  & slopes\_online\_mean & dist\_deepest          & 6.7 m          & 0.32          & 66 \%          \\
            slope\_mean          & slope\_mean          & dists\_deepest         & 6.8 m          & 0.36          & 59 \%          \\
            slope\_pnt           & slopes\_online\_mean & dists\_deepest         & 6.8 m          & 0.25          & 73 \%          \\
            slope\_pnts          & slope\_online\_mean  & dist\_deepest          & 6.9 m          & 0.3           & 71 \%          \\
            slope\_online\_mean  & slope\_online\_mean  & dist\_deepest          & 6.9 m          & 0.32          & 68 \%          \\
            slope\_online\_mean  & slope\_online\_mean  & dists\_deepest         & 6.9 m          & 0.33          & 65 \%          \\
            slope\_mean          & slopes\_online\_mean & dists\_deepest         & 7 m            & 0.24          & 65 \%          \\
            slope\_mean          & slope\_mean          & dist\_deepest          & 7.1 m          & 0.4           & 64 \%          \\
            slopes\_online\_mean & slope\_mean          & dist\_deepest          & 7.1 m          & 0.37          & 56 \%          \\
            slope\_mean          & slope\_online\_mean  & dist\_deepest          & 7.1 m          & 0.3           & 69 \%          \\
            slopes\_online\_mean & slopes\_online\_mean & dists\_deepest         & 7.2 m          & 0.32          & 63 \%          \\
            slopes\_online\_mean & slopes\_online\_mean & dist\_deepest          & 7.3 m          & 0.25          & 64 \%          \\
            slope\_pnt           & slope\_mean          & dists\_deepest         & 7.3 m          & 0.35          & 61 \%          \\
            slopes\_online\_mean & slope\_mean          & dists\_deepest         & 7.3 m          & 0.36          & 60 \%          \\
            slope\_online\_mean  & slopes\_online\_mean & dists\_deepest         & 7.3 m          & 0.29          & 58 \%          \\
            slope\_pnts          & slope\_online\_mean  & dists\_deepest         & 7.4 m          & 0.27          & 64 \%          \\
            slopes\_online\_mean & slope\_online\_mean  & dists\_deepest         & 7.4 m          & 0.33          & 67 \%          \\
            slopes\_online\_mean & slope\_online\_mean  & dist\_deepest          & 7.5 m          & 0.26          & 61 \%          \\
            slope\_pnt           & slopes\_online\_mean & dist\_deepest          & 7.5 m          & 0.33          & 69 \%          \\
            slope\_mean          & slopes\_online\_mean & dist\_deepest          & 7.6 m          & 0.26          & 64 \%          \\
            slope\_pnt           & slope\_online\_mean  & dist\_deepest          & 7.7 m          & 0.27          & 68 \%          \\
            slope\_pnts          & slopes\_online\_mean & dists\_deepest         & 7.8 m          & 0.3           & 65 \%          \\
            slope\_pnt           & slope\_online\_mean  & dists\_deepest         & 7.9 m          & 0.27          & 67 \%          \\
            slope\_mean          & slope\_online\_mean  & dists\_deepest         & 7.9 m          & 0.31          & 60 \%          \\
            \hline
      \end{tabular}
\end{table}
